---
title: "ALPPACA ANI Report"
date: "`r Sys.Date()`"
output:
  html_document:
    mathjax: null
    code_folding: "hide"
    toc: true
    toc_float:
      collapsed: false
params:
  ani_data:
    value: x
---
# Run information

Date generated: `r Sys.Date()`

## R Packages
These are the R packages used to generate this report.
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(ggplot2)
library(patchwork)
library(readr)
library(formattable)
library(kableExtra)
library(ggdist)

packages <- (.packages())
clean_pkg_version <- function(x) {
  y <- getNamespaceVersion(x)
  
  z <- data.frame("Package" = x,
                  "Version" = y)
  
  row.names(z) <- NULL
  
  return(z)
}
lapply(packages, function(x) clean_pkg_version(x)) %>%
  bind_rows() %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

<br>

## Tool versions
These are the tools used to run the analysis within the ani workflow of ALPPACA.
```{r, message=FALSE, warning=FALSE}
tool_data <- data.frame(
  Tool = c("FastANI"),
  Version = c("1.33")
)
tool_data %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

<br>

<br>

# Tool information

## FastANI
The figure below presents the summarised average nucleotide identity (ANI)
density for all included isolates. The black diamond and the bold line 
in the boxplot represent the median value. Vertical blue lines represent
each data point. The blue area above the horizontal line represent the
density of samples for each ANI value. 

```{r, dpi=600, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
# helper Sorting order to remove duplicates
keysort <- function(x,y) {
  sorted <- sort(c(x,y))
  paste0(sorted[1], "_", sorted[2])
}

# Removes paths and extension from ID
regex_pattern <- regex(
  "^(.*/)|([:punct:]?.((fasta)|(fna)|(fa)))"
  )

file_list <- list.files(pattern = "_fastani.txt")

fastani_data <- lapply(file_list, function(x) {
  read_delim(
    x, 
    delim = "\t",
    col_names = FALSE
    )
  }
) %>% 
  bind_rows() %>%
  rename("ID1" = X1,
         "ID2" = X2,
         "ANI" = X3,
         "Count Bidirectional Fragments Mappings" = X4,
         "Total Query Fragments" = X5) %>%
  dplyr::mutate(ID1 = str_remove_all(ID1, regex_pattern),
                ID2 = str_remove_all(ID2,  regex_pattern),
                ANI = as.numeric(ANI),
                sortkey =  purrr::map2(ID1, ID2, ~keysort(.x, .y))) %>%
  # removal self comparison
  filter(ID1 != ID2) %>%
  arrange(sortkey) %>%
  group_by(sortkey) %>%
  # keep first one of comparison (ID1,ID2) == (ID2,ID1)
  slice(1) %>%
  ungroup() %>%
  select(-sortkey)

write.table(fastani_data,
            "FASTANI_cleaned.txt",
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

ggplot(fastani_data, aes(x = ANI, y = 1)) +
  geom_point(
    aes(y = 1 - 0.03),
    shape = "|",
    color = "#1f78b4",
    size = 8,
    alpha = 0.33
  ) +
  stat_boxplot(
    aes(y = 1 - 0.09),
    geom = "errorbar",
    width = 0.05
  ) +
  geom_boxplot(
    aes(y = 1 - 0.09),
    width = 0.05,
    fill = "#a6cee3"
  ) +
  stat_halfeye(
    shape = 18,
    point_size = 4,
    interval_size = 1.8,
    adjust = 0.5,
    trim = TRUE,
    height = 0.8,
    .width = c(0,1),
    fill = "#a6cee3"
  ) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 14),
	axis.title.x = element_text(size = 16))
```

