---
title: "ALPPACA CGMLST Report"
date: "`r Sys.Date()`"
output:
  html_document:
    mathjax: null
    code_folding: "hide"
    toc: true
    toc_float:
      collapsed: false
params:
  allelecall_data:
    value: x
  schema_summary:
    value: x
  loci_summary:
    value: x
  result_stats:
    value: x
  hamming_dists:
    value: x
  dendrogram:
    value: x
---

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(readr)
library(ggtree)
library(plotly)
library(kableExtra)
```

---

# Schema information
These tables contain information about the schema used in the cgMLST analysis.
```{r, message=FALSE, warning=FALSE}
schema_data <- read_delim(
  params$schema_summary,
  delim = "\t",
  col_names = TRUE
)

excluded_genes <- schema_data %>%
  filter(Number_representatives == 0)

excluded_genes %>%
  kbl(caption = "Number of invalid genes in the schema.") %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)

```

The schema had a total of `r nrow(schema_data)` genes. Of these, `r nrow(excluded_genes)` genes were excluded. 


```{r, message=FALSE, warning=FALSE, dpi=600, fig.width=8, fig.height=5, fig.cap="**Figure 1**: Schema loci summary stats.", fig.align='center'}
loci_data <- read_delim(
  params$loci_summary,
  delim = "\t",
  col_names = TRUE
) %>%
  pivot_longer(cols = -Locus,
               names_to = "Type",
               values_to = "n") %>%
  mutate(Type = factor(Type,
                       levels = c(
                         "EXC",
                         "INF",
                         "LNF",
                         "PLNF",
                         "PLOT3",
                         "PLOT5",
                         "LOTSC",
                         "NIPH",
                         "NIPHEM",
                         "PAMA",
                         "ALM",
                         "ASM",
                         "Total_CDS"
                       )))

palette_ids <- unique(loci_data$Type)
palette <- RColorBrewer::brewer.pal(name = "Paired", n = 12)
names(palette) <- palette_ids

ggplot(loci_data, aes(Type, n, fill = Type)) +
  stat_boxplot(geom = "errorbar",
               width = 0.5) +
  geom_boxplot() +
  scale_fill_manual(values = palette) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid = element_blank())
```



<br>

---

# AlleleCalling stats

This section presents the AlleleCalling results and statistics. 

```{r, message=FALSE, warning=FALSE, fig.height=5, fig.width=8, dpi=600, fig.cap="**Figure 2**: AlleleCall result overview.", fig.align='center'}
allelecalling_data <- read_delim(
  params$allelecall_data,
  delim = "\t",
  col_names = TRUE
) %>%
  pivot_longer(cols = -FILE,
               names_to = "Type",
               values_to = "n") %>%
  mutate(Type = factor(Type,
                       levels = c(
                         "EXC",
                         "INF",
                         "LNF",
                         "PLNF",
                         "PLOT3",
                         "PLOT5",
                         "LOTSC",
                         "NIPH",
                         "NIPHEM",
                         "PAMA",
                         "ALM",
                         "ASM"
                       )))

ggplot(allelecalling_data, aes(Type, n, fill = Type)) +
  stat_boxplot(geom = "errorbar",
               width = 0.5) +
  geom_boxplot() +
  scale_fill_manual(values = palette) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid = element_blank())

```
<br>

```{r, message=FALSE, warning=FALSE, dpi=600, fig.height=8, fig.width=8, fig.cap="**Figure 3**: Number of compared and missing alleles per isolate.", fig.align='center'}
allele_data <- read_delim(
  params$hamming_dists,
  delim = "\t",
  col_names = TRUE
)

compared_alleles <- ggplot(allele_data, aes(isolate1, compared_alleles)) +
  stat_boxplot(geom = "errorbar", width = 0.3) +
  geom_boxplot(fill = "grey70", width = 0.5) +
  labs(y = "N Compared Alleles") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank())

missing_alleles <- ggplot(allele_data, aes(isolate1, missing_alleles)) +
  stat_boxplot(geom = "errorbar", width = 0.3) +
  geom_boxplot(fill = "grey70", width = 0.5) +
  labs(x = "Isolates",
       y = "N Missing Alleles") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

compared_alleles / missing_alleles
```


<br>

---

# Distance calculations

```{r, message=FALSE, warning=FALSE, dpi=600, fig.height=5, fig.width=8, fig.cap="**Figure 4**: Hamming distance and number of compared and missing alleles.", fig.align='center'}
hamming_distances <- read_delim(
  params$hamming_dists,
  delim = "\t",
  col_names = TRUE
) %>%
  rename("Hamming" = hamming,
         "Compared alleles" = compared_alleles,
         "Missing alleles" = missing_alleles) %>%
  pivot_longer(cols = -c(isolate1, isolate2),
               names_to = "Type",
               values_to = "value") %>%
  mutate(Type = factor(Type,
                       levels = c("Hamming",
                                  "Compared alleles",
                                  "Missing alleles")))

p1 <- hamming_distances %>%
  filter(Type == "Hamming") %>%
  ggplot(aes(Type, value)) +
  geom_violin(trim = FALSE,
              scale = "count",
              fill = "#fdb462") +
  stat_boxplot(geom = "errorbar", width = 0.05) +
  geom_boxplot(width = 0.05) +
  geom_jitter(width = 0.2,
              alpha = 0.2) +
  labs(y = "Hamming Distance") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())


p2 <- hamming_distances %>%
  filter(Type != "Hamming") %>%
  ggplot(aes(Type, value, fill = Type)) +
  geom_violin(trim = FALSE,
              scale = "count") +
  stat_boxplot(geom = "errorbar", width = 0.05) +
  geom_boxplot(width = 0.05) +
  geom_jitter(width = 0.2,
              alpha = 0.2) +
  scale_fill_manual(values = c("Compared alleles" = "#80b1d3",
                               "Missing alleles" = "#fb8072")) +
  labs(y = "N") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")


p1 + p2 +
  plot_layout(widths = c(0.5, 1))


```

<br>

---

# Clustering

```{r, warning=FALSE, message=FALSE, fig.cap="**Figure 5**: Dendrogram based on the selected clustering method.", fig.align='center'}
tree <- read.tree(
  params$dendrogram
)

tree_plot <- ggtree(tree) +
  geom_text(aes(label = label),
            hjust = 0,
	    size = 4) +
  geom_treescale()

ggplotly(tree_plot) %>%
  plotly::style(textposition = "right")
```

---

<details>
  <summary>Tool and package versions</summary>

Date generated: `r Sys.Date()`

**R Packages**

These are the R packages used to generate this report.
```{r, message=FALSE, warning=FALSE}
packages <- (.packages())
clean_pkg_version <- function(x) {
  y <- getNamespaceVersion(x)
  
  z <- data.frame("Package" = x,
                  "Version" = y)
  
  row.names(z) <- NULL
  
  return(z)
}

lapply(packages, function(x) clean_pkg_version(x)) %>%
  bind_rows() %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

**Tool versions**

These are the tools used to run the analysis within the cgMLST workflow of ALPPACA.
```{r, message=FALSE, warning=FALSE}
tool_data <- data.frame(
  Tool = c("ChewBBACA","R"),
  Version = c("3.1.2","4.1.2")
)
tool_data %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

</details>

