---
title: "ALPPACA Core Gene Report"
date: "`r Sys.Date()`"
output:
  html_document:
    mathjax: null
    code_folding: "hide"
    toc: true
    toc_float:
      collapsed: false
params:
  pangenome_data:
    value: x
  snpdist_report:
    value: x
  phylo_tree:
    value: x
  phylo_data:
    value: x
---
```{r, include=FALSE, message=FALSE, warning=FALSE}
library(stringr)
library(readr)
library(dplyr)
library(tidyr)
library(rlang)
library(ggtree)
library(plotly)
library(patchwork)
library(ggplot2)
library(viridis)
library(formattable)
library(kableExtra)
```


# Pangenome analysis
<center>*Number of genes detected in the pangenome analysis. The core genes were aligned and used in the phylogenetic analysis.*</center>
```{r}
read_delim(params$pangenome_data,
           delim = "\t",
           show_col_types = FALSE,
           col_names = FALSE) %>%
  rename("Type" = X1,
         "Definition" = X2,
         "Number of genes" = X3) %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

<br>

# SNPdist
<center>*SNP distance summary statistics for all isolates.*</center>
```{r, message=FALSE, warning=FALSE}
snpdist_data <- read_delim(params$snpdist_report,
                           delim = "\t",
                           show_col_types = FALSE) %>%
  rename("isol1" = 1) %>%
  pivot_longer(cols = -isol1,
               names_to = "isol2",
               values_to = "value")
snpdist_data %>%
  filter(isol1 != isol2) %>%
  mutate(Median = median(value),
         Mean = round(mean(value), 0),
         Range = paste0(min(value), " - ", max(value))) %>%
  select(Mean, Median, Range) %>%
  head(1) %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```


```{r, message=FALSE, warning=FALSE, dpi=600, fig.cap="**Figure 3**: SNP distance matrix of all included isolates. A lighter color represents a smaller SNP distance.", fig.align='center'}
ggplot(snpdist_data, aes(isol1, isol2, fill = value)) +
  geom_tile() +
  labs(fill = "SNP distance") +
  scale_fill_viridis(direction = -1,
                     option = "D") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
	axis.text = element_text(size = 4),
        panel.grid = element_blank(),
        axis.title = element_blank()) +
  coord_fixed()
```

<br>

# IQTree

```{r, message=FALSE, warning=FALSE}
# Read IQTree data
iqtree_data <- readLines(params$phylo_data)
aln_info <- sub(
  "Input data: ",
  "",
  iqtree_data[str_detect(iqtree_data,"Input data:")]
)
const_sites <- sub(
  ".+= (.+) of all sites)",
  "\\1",
  iqtree_data[str_detect(iqtree_data,"Number of constant sites:")]
)
parsimony_sites <- sub(
  "Number of parsimony informative sites: ",
  "",
  iqtree_data[str_detect(iqtree_data,"Number of parsimony informative sites:")]
)
iqtree_model <- sub(
  "Model of substitution: ",
  "",
  iqtree_data[str_detect(iqtree_data,"Model of substitution: ")]
)
iqtree_cpu <- sub(
  "Total CPU time used: ",
  "",
  iqtree_data[str_detect(iqtree_data,"Total CPU time used:")]
)
iqtree_wallclock <- sub(
  "Total wall-clock time used: ",
  "",
  iqtree_data[str_detect(iqtree_data,"Total wall-clock time used:")]
)
```

IQTree was run on an alignment composed of `r aln_info`, where `r const_sites` of the alignment were constant sites. The number of parsimonious sites were `r parsimony_sites`. IQTree detected the evolutionary model `r iqtree_model`. IQTree used `r iqtree_cpu` of CPU time, which converts to `r iqtree_wallclock` of wall-clock time.

<br>

<br>

# Phylogenetic tree

Here, the phylogenetic tree generated by IQTree is plotted. The figure is interactive, and you can zoom in and out and pan as you see fit.

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=8, fig.cap="**Figure 4**: Maximum likelihood phylogenetic tree generated by IQTree. Bootstrap values are presented for each node. The tree is unrooted", fig.align='center'}
tree <- read.tree(params$phylo_tree)
tree_plot <- ggtree(tree) +
  geom_text(aes(label = label),
            hjust = 0,
	    size = 2) +
  geom_treescale()
ggplotly(tree_plot) %>%
  plotly::style(textposition = "right")
```

---

<details>
  <summary>Tool and package versions</summary>

Date generated: `r Sys.Date()`

**R Packages**

These are the R packages used to generate this report.
```{r, message=FALSE, warning=FALSE}
packages <- (.packages())
clean_pkg_version <- function(x) {
  y <- getNamespaceVersion(x)
  
  z <- data.frame("Package" = x,
                  "Version" = y)
  
  row.names(z) <- NULL
  
  return(z)
}
lapply(packages, function(x) clean_pkg_version(x)) %>%
  bind_rows() %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```


**Tool versions**

These are the tools used to run the analysis within the core genome workflow of ALPPACA.
```{r, message=FALSE, warning=FALSE}
tool_data <- data.frame(
  Tool = c("Bakta",
           "Panaroo",
           "Snp-dists",
           "Snp-sites",
           "IQTree"),
  Version = c("1.9.1",
              "1.3.4",
              "0.8.2",
              "2.5.1",
              "2.2.6")
)
tool_data %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "responsive"),
                full_width = TRUE)
```

</details>
